\textcolor{blue}{Local maximum entropy material point method applied to quasi-brittle fracture.}

### 2020-10-06 2020_EFM_MPM_Eigensoftening.tex
Revisión MMP

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{lineno}
% \usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx} % figuras
% \usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{multirow} 
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However, the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical
\textcolor{red}{artifices}  have been proposed with the aim of reproducing such a
complex behaviour. These techniques vary from the employment of cohesive
approaches~\cite{Barenblatt,Hilleborg_1976}, or to the gradual insertion of
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries. To alleviate these shortcomings of the
\acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018}, the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}, and more
recently improving it capability to predict the crack path in 3D cases
\cite{YUN_2019}.

The simulation of fracture propagation in a more accurate and effective
way can be considered as one of the ongoing goals of the development of novel spatial discretisation methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
\acrfull{efgm} \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013}, the
\acrfull{sph} \cite{Wang_2020,Wang_2019}, the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others. \\

Concerning \acrshort{mpm}, fracture has been treated numerically through two different ways. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003,Nairn_2006}. The basis of this methodology consist of the removal of the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analysed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the stress
intensity factor or the J-integral discussed by Guo \& Nairn
(2004)\cite{Nairn_2004}. 
On the other hand, the foundation of the second approach is the introduction of failed
material points to describe the crack evolution. In these methods, the
formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Chen {\it et al.}\cite{Zhenmao_2005} and
Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et
  al.} \cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}.\\

Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
crack sets are approximated by means of eigen-deformations, which
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle fracture
was carried out by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique, the concept of ``erosion'' of
the material point is depicted, being the material point able to remain intact or become completely failed or eroded, when no loading
capacity is bore. This method has been successfully applied to simulate high
complex phenomena such as dynamic fragmentation of metals
\cite{Li_2015}. Recently, Zhang {\it et al.}
\cite{Zhang_EE_2020} adopted the eigenerosion to resolve the dynamic
fracture of brittle materials in the \acrshort{mpm}
framework. Nonetheless, when quasi-brittle materials are simulated
though this approach, the results exhibit an overestimation of tensile
stress and strain peaks. Furthermore, since the work of Zhang {\it et al.}
\cite{Zhang_EE_2020} is made with conventional \acrshort{mpm}/\acrshort{gimp}, stress oscillations are to be faced.\\

To overcome the limitations observed in the EigenMPM
\cite{Zhang_EE_2020}, the present research proposes the eigensoftening
algorithm developed by Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} for the \acrshort{otm} framework and
engineered for quasi-brittle materials. Inspired in the concept of the
crack band model \cite{Bazant83}, since energy dissipation is performed through the softened (or failed) volume, this methodology is able to capture the gradual rather than abrupt
dissipation of the fracture energy. Moreover, in order to mitigate stress
oscillations as well as to deal with tensile instabilities, such those that
occurs when particles crosses an element boundary, the work of Molinos~\etal~\cite{Molinos2020} is followed. The traditional \acrshort{mpm} technique is enhanced through the \acrfull{lme} approximation technique
\cite{Arroyo2006} for the spatial discretisation as well as a explicitly \acrshort{mpm} \acrfull{npc} scheme is derived for the time discretisation.\\

@@

Since the \acrshort{mpm} possesses the advantages of both Lagrangian and Eulerian descriptions, no  element distortion takes place in the \acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
\textcolor{blue}{the explicit predictor-corrector time integration scheme proposed by Molinos~\etal~\cite{Molinos2020} is adopted}. The
purpose of this choice is motivated due to its proved robustness and
stability for dynamic computations. In the first stage, the nodal
velocity predictor \textcolor{red}{$ \vec{v}_{I}^{pred}$} is computed through
\textcolor{red}{particles} velocities \textcolor{red}{$\vec{v}_p$} and accelerations \textcolor{red}{$\vec{a}_p$} and evaluated
in the $k$ time step,
%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
 \label{eq:Predictor-velocity}
    \vec{v}_{I}^{pred} =  \textcolor{blue}{\sum_{p\ \in\ \mathcal{C}}} \frac{ N_{I}(\vec{x}_p^k )\ m_p\ (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and \textcolor{red}{$N_{I}(\vec{x}_p^k) $} is the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}}
          \frac{N_{I}(\vec{x}_p^k)\ \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
        \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n +  \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
         N_{I}(\vec{x}_p^k)\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

\subsection{Spatial interpolation technique : \acrlong{lme} approximants}
\label{sec:2.2}
\acrfull{lme} approximation scheme was introduced by Arroyo \& Ortiz
(2006)\cite{Arroyo2006} as a bridge between finite elements and
meshfree methods. The key idea under this interpolation technique is
the interpretation of each nodal value $N_I$ as a
probability. Related with this definition, two important limits are
introduced. First the maximum-entropy (\textit{max-ent}) limit, which
ensures a \textit{unbiased statistical inference} based on the nodal
data as states the Jayne's\cite{Jaynes1957} principle of
\textit{maximum entropy}. And second the Delaunay triangulation that
warranties the \textit{least width} shape function support. To reach
a compromise between two competing objectives, a Pareto set is

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Zhenmao_2005,Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. The
  evaluation of both approaches \cite{Navas_2017_ES} against
  experimental measures show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. Furthermore,
the accuracy of this algorithm has been proven for complex fracture
patterns such as the ones obtained with fiber \textcolor{red}{reinforced} concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

Implementation details can be consulted in \ref{sec:eigens-algor-1}.

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology. Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms like Crossing Number \cite{Shimrat_1962} or Correct
Even-Odd \cite{Galetzka_et_al_2017}, the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead of with the belonging element as heretofore has been
done.
Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest nodes are defined, the particle
possesses automatically a list of \textit{tributary nodes}. This has two
immediate benefits: the first one is that shape functions with global
support as \acrshort{lme} are easier to be implemented,
and the second one is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, it is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The second application to be considered analyses the bending behavior of semicircular specimens of Johnstone rocks, text which were experimentally carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions. This capability was assessed previously within a
\acrshort{otm} implementation \cite{Navas_2018_ES}.

@@

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Finally, further improvements of the present research are exposed. The
algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced
concretes\cite{Navas_2018_ES,Ruiz2019}. In addition, the eigensoftening
approach can be extrapolated to rock fracture in the same manner the
eigenerosion was~\cite{Wang2017}, taking into account that the
volumetric stiffness remains in order to support bulk loads. Also the
benefits of \acrshort{lme} can be exploited by adapting it through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would
enhance the localisation properties of the proposed algorithms in more
sophisticated domains. And the implementation of a contact algorithm to model other impact cases such as bullets collisions.

>>>>

\documentclass[preprint,12pt,a4paper]{elsarticle}
\usepackage{lineno}
% \usepackage{hyperref}
\usepackage{float}
\usepackage{graphicx} % figuras
% \usepackage{subfig}
\usepackage{subfigure} % subfiguras
\usepackage{multirow} 
\usepackage{color}
\usepackage{soul}
\usepackage{cancel}
\usepackage{textcomp}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
The failure mechanisms of quasi-brittle materials and the  crack propagation have been studied from different fields such as materials science, structural engineering, earth science and petroleum engineering . The prediction of crack propagation has been approached with different numerical techniques. The \acrfull{fem} is widely used in industrial application. However, the presence of cracks is a violation of the continuity requirement of the \acrshort{fem}. To overcome this issue, numerous numerical \textcolor{red}{artifices} have been proposed with the aim of reproducing such a
complex behaviour. \textcolor{blue}{An example of this}, is the gradual insertion of
cohesive elements \cite{Ortiz_1999} at solid elements boundaries. To alleviate these shortcomings of the \acrshort{fem}, the \acrfull{xfem} in conjunction with level set
technique was introduced by Belytschko {\it et al.}
\cite{Belytschko_03} to handle arbitrary crack paths level
set. Some relevant contributions to the \acrshort{xfem} are the
modeling micro-scale crack propagation \cite{HUANG_2018} \textcolor{blue}{or} the
introduction of a cohesive zone model \cite{NAGASHIMA_2016}.\\
The simulation of fracture propagation in a more accurate and effective
way can be considered as one of the ongoing goals of the development of novel spatial discretisation methods like meshfree techniques. Some
examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Nairn_2003,NAIRN_2018}, the
\acrfull{efgm}~\cite{BELYTSCHKO_1995b,BELYTSCHKO_2000}, the
\acrfull{sph}~\cite{Wang_2020,Wang_2019}, the \acrfull{otm}~\cite{Pandolfi_2013,Li_2015} or Peridynamics~\cite{HA_2011,RABCZUK_2017} among others.\\
Concerning \acrshort{mpm}, fracture has been treated numerically through two different ways. On the one hand, the ``CRAcks with Material Points (CRAMP)'' is proposed~\cite{Nairn_2003}. The basis of this methodology consist of the removal of the restriction
of the single-valued velocity field close to the crack through, at least, two sets of nodes. In this method, different labels are assigned to the material points
and nodes to distinguish if they are in the same side of the crack or
not. Under this approach, crack surface is described with line segments
in 2D and triangle patches in 3D cases. The chosen criteria for crack
propagation is based on such parameters as energy release rate
analysed by Tan \& Nairn (2002)\cite{Nairn_2002}. On the other hand, the foundation of the second approach is the introduction of failed material points to describe the crack evolution. In these methods, the formation of failed points describes the nucleation of cracks, and
thereafter its propagation and branching. Consequently, the position
of the crack does not need to be explicitly stated. This fact represents
significant advantages over the ``CRAMP''. Under this approach, the
prediction of failure evolution is computed with a decohesion model,
which has been discussed by Schreyer {\it et al.}\cite{Schreyer_2002}. Successful simulations are provided within the framework of the fracture of brittle materials by Chen {\it et al.} \cite{Chen_2002}.\\

Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
crack sets are approximated by means of eigen-deformations, which
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle fracture
was carried out by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique, the concept of ``erosion'' of
the material point is depicted, being the material point able to remain intact or become completely failed or eroded, when no loading
capacity is bore. This method has been successfully applied to simulate high
complex phenomena such as dynamic fragmentation of metals
\cite{Li_2015}. Recently, Zhang {\it et al.}
\cite{Zhang_EE_2020} adopted the eigenerosion to resolve the dynamic
fracture of brittle materials in the \acrshort{mpm}
framework. Nonetheless, when quasi-brittle materials are simulated
though this approach, the results exhibit an overestimation of tensile
stress and strain peaks. Furthermore, since the \textcolor{red}{aforementioned} work is made with conventional \acrshort{mpm}/\acrshort{gimp}, stress oscillations are to be faced.\\

To overcome the limitations observed in the EigenMPM, the present research proposes the eigensoftening algorithm developed by Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} for the \acrshort{otm} framework and
engineered for quasi-brittle materials. Inspired in the concept of the
crack band model \cite{Bazant83}, since energy dissipation is performed through the softened (or failed) volume, this methodology is able to capture the gradual rather than abrupt
dissipation of the fracture energy. Moreover, in order to mitigate stress
oscillations as well as to deal with tensile instabilities, such those that
occurs when particles crosses an element boundary, the work of Molinos~\etal~\cite{Molinos2020} is followed. The traditional \acrshort{mpm} technique is enhanced through the \acrfull{lme} approximation technique \cite{Arroyo2006} for the spatial discretisation as well as a explicitly \acrshort{mpm} \acrfull{npc} scheme is derived for the time discretisation.\\

@@

Since the \acrshort{mpm} possesses the advantages of both Lagrangian and Eulerian descriptions, no  element distortion takes place in the \acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation in a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research,
\textcolor{blue}{the explicit predictor-corrector time integration scheme proposed by Molinos {\it et al.} \cite{Molinos2020} has been adopted. The purpose of this choice is motivated due to its proved robustness and stability for dynamic computations, Navas {\it et al.} \cite{Navas2018a}}. In the first stage, the nodal velocity predictor \textcolor{red}{$ \vec{v}_{I}^{pred}$} is computed through
\textcolor{red}{particles} velocities \textcolor{red}{$\vec{v}_p$} and accelerations \textcolor{red}{$\vec{a}_p$} and evaluated
in the $k$ time step,
%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
 \label{eq:Predictor-velocity}
    \vec{v}_{I}^{pred} =  \textcolor{blue}{\sum_{p\ \in\ \mathcal{C}}} \frac{ N_{I}(\vec{x}_p^k )\ m_p\ (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and \textcolor{red}{$N_{I}(\vec{x}_p^k) $} is the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}}
          \frac{N_{I}(\vec{x}_p^k)\ \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
        \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n +  \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
         N_{I}(\vec{x}_p^k)\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}
The complete pseudo-algorithm it is summarised in \ref{sec:expl-pred-corr}.

\subsection{Spatial interpolation technique : \acrlong{lme} approximants}
\label{sec:2.2}
\acrfull{lme} approximation scheme was introduced by Arroyo \& Ortiz
(2006)\cite{Arroyo2006} as a bridge between finite elements and
meshfree methods. \textcolor{blue}{The basic idea of the shape functions based on such an estimate is to interpret the shape function $N_I(\vec{x})$ as a probability. This allows us to introduce two important limits: the principle of maximum-entropy (\textit{max-ent}) statistical
inference stated by Jaynes \cite{Jaynes1957}, and the Delaunay triangulation
which ensures the minimal width of the shape function.}

\textcolor{blue}{This approximation scheme represents an optimal compromise, in the sense of
Pareto, between the \textit{unbiased statistical inference} based on
the nodal data which leads to the principle of \textit{Maximum-Entropy}
stated by Jaynes \cite{Jaynes1957}, and the definition of local shape
functions of \textit{least width} the least biased shape functions.
Following \cite{Arroyo2006}, entropy of a discrete random variable can be defined as the uncertainty of the random variable. A measure of this uncertainty can be obtained by means of Shannon\textquotesingle s entropy:}
  \label{eq:Shannon-entropy}
  \textcolor{blue}{H(p_1,\ldots,p_n) = -\sum^{N_n}_{I=1}{p_I\ \log p_I }}
\textcolor{blue}{where $p_I$ stands for probability of the random variable outcomes. By interpreting these probabilities as the shape functions $N_I(\vec{x})$ of an approximation scheme, equation \eqref{eq:Shannon-entropy} can be regarded as a measure of the uncertainty of the approximation.} To reach a compromise between two competing objectives, a Pareto set is

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be
modelled by failing particles according to a suitable failure
criterion. When material points are failed, they are assumed to have
null stress tensor. To reproduce this behaviour in the present
research, the eigensoftening algorithm is introduced in the 
\acrshort{mpm} framework as an alternative approach to the decohesion
model \cite{Schreyer_2002}. The eigensoftening concept
was originally developed by Navas {\it et al.}
(2017)\cite{Navas_2017_ES} as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. The
  evaluation of both approaches \cite{Navas_2017_ES} against
  experimental measures show that the eigenerosion algorithm significantly overestimates the tensile stress and the strain peaks, meanwhile eigensoftening captures the forces and crack
patterns accurately. Furthermore, the accuracy of this algorithm has been proven for \textcolor{blue}{mixed-mode} fracture patterns such as the ones obtained with fiber \textcolor{red}{reinforced} concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

Implementation details can be consulted in \ref{sec:eigens-algor-1}. \textcolor{blue}{Unfortunately, this method is not free of pitfalls. The first one involves the isotropic model of failure assumption of the eigensofteing, therefore, complex fracture patterns such as delamination can not be simulated. A possible solution for this incovenience is the use of anisotropic damage model. Regarding this, the current work can be extended following the approach proposed by Nairn~\cite{NAIRN_2018}. The second one is handling crack contact. After particle fails, the stress tensor of it is enforced to be zero in all the directions, of course, this is a direct consequence of the isotropic model of failure assumption. Consequently, a spurious slip condition is enforced in the tangent direction of the crack, also the material loose its toughness to compression in the fracture interface. This limitation was noticed earlier by in Pandolfy and Ortiz \cite{Pandolfi_2012} two alternatives to overcome this limitation where briefly discussed. Again, the use of an anisotropic damage model~\cite{NAIRN_2018} would overcome this limitation.}

\subsection{$\epsilon$-neighbourhood reconstruction : A node-linked method}
\label{sec:epsil-neighb-reconst}
The construction of the $\epsilon$-neighbourhood is a major issue of the proposed methodology. Since this operation could be extremely demanding, optimal
numerical implementation should be employed. In opposite with the Cell-linked
method proposed by Allen \& Tildesley \cite{Allen_et_al_1989} where
the definition of a numerical tolerance is required for auxiliar
algorithms, the proposed methodology adopted herein is linked to the shape function neighbourhood. In order to avoid the aforementioned error prone
techniques, we endeavour to exploit the meshfree benefits of
the \acrshort{lme} approximants to introduce a node-linked method. Due
to the global support of this interpolation technique, the classical
interpretation of the \acrshort{fem} element typically adopted in
\acrshort{mpm} simulations is not required anymore. Instead, a linked
list with the surrounding nodes for each node is defined at the initial
time. Also, each particle has to be associated with the
closest node instead of with the belonging element as heretofore has been
done.
Now lets define the concept of a particle \textit{tributary nodes} as the
list of nodes close to it among which the information transfers
occurs. Once the closest nodes are defined, the particle
possesses automatically a list of \textit{tributary nodes}. This has two
immediate benefits: the first one is that shape functions with global
support as \acrshort{lme} are easier to be implemented,
and the second one is the increase of the efficiency of the $\epsilon$-neighbourhood
reconstruction. Regarding this last benefit, since the current
particle has a \textit{tributary nodes} list assigned to it, and
consequently each node has a linked list with a set of
\textit{tributary particles}\footnote{Notice that the list of the
  particles close to each node is unique for each node and is updated
  in each time step.} close to it, it is possible to define efficiently a
search list with those particles close to the current particle.
In the end, particles out of the area of influence, Figure
\ref{fig:Failed-particles}, are pop out of the final
$\epsilon$-neighbourhood. A final remark about this new approach is
its ability to reconstruct easily a search list with the support of
the initial connectivity of the mesh. It provides a suitable basis for
an easy implementation of other interpolation techniques with minimum
coding effort. A detailed explanation of the proposed algorithm can be
found in \ref{sec:node-linked-method-1}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The second application to be considered analyses the bending behavior of semicircular specimens of Johnstone rocks, text which were experimentally carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions.

@@

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textcolor{blue}{Next, further improvements of the present research are exposed. Firstly, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads. Secondly, the algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes by means of empirical definitions of $\chi$ curve~\cite{Navas_2018_ES}. However, capturing fracture patters in anisotropic media such as carbon fibre plates requires the redefinition of the approximation towards an anisotropic softening damage model. In addition, this improvement also would allows the model to handle crack contact. Finally, the implementation of a contact algorithm to model other impact cases such as bullets collisions.}
